---
title: "geocache"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In a way, this is a silly R package.  It's a wrapper around the very useful [rappdirs package](https://CRAN.R-project.org/package=rappdirs), but with a some extra tools that are helpful for use with the [geodata package](https://CRAN.R-project.org/package=geodata)

### Requirements

+ [R v4.1+](https://www.r-project.org/)

From CRAN

+ [rappdirs package](https://CRAN.R-project.org/package=rappdirs)

### Installation

```
remotes::install_github("BigelowLab/geocache")
```

### Establish a initial directory for geodata

You can use the [geodata package](https://CRAN.R-project.org/package=geodata) in such a way that you request a fresh download each time establish a new R session. As an alternative, you could download and save a local copy for retrieval and use in subsequent sessions.  There is a balance between saving too much data locally and the convenience of locally stored data; each of us has to find that balance. 

Using [rappdirs](https://CRAN.R-project.org/package=rappdirs) allows you to define either a personal data directory or a site-wide directory.  See `?rappdirs` for more details.  I'm going to choose the default personal directory.

>Note that you can choose an entirely different path unrelated to the ones provided by [rappdirs](https://CRAN.R-project.org/package=rappdirs).  In that case, just save the path description in a hidden text file, `~/.geocache`, and things will work seamlessly.

```{r first use}
library(geocache)
path <- geocache::geocache_path()
path
```

### Now get some geodata

The [geodata](https://CRAN.R-project.org/package=geodata) serves up a wide variety of point, vector and raster. Let's use the example of future climate projections from [CMIP6](https://esgf-node.llnl.gov/projects/cmip6/)

Since [geodata](https://CRAN.R-project.org/package=geodata) provides a wide variety of data it might be good to organize the geocache with subdirectories.  Let's start with one for `cmip6`.

```{r subdirectory, warn = FALSE}
cmip6_path <- file.path(path, 'cmip6')
if (!dir.exists(cmip6_path)) ok <- dir.create(cimp6_path)
```

Now let's get some data and store it there...

```{r get cmip6}
x2021 <- geodata::cmip6_world(model = "CNRM-CM6-1", ssp = "585", 
                          res = 10, time = "2021-2040",
                          var = "bioc",
                          path = cmip6_path)

x2041 <- geodata::cmip6_world(model = "CNRM-CM6-1", ssp = "585", 
                          res = 10, time = "2041-2060",
                          var = "bioc",
                          path = cmip6_path)

x2061 <- geodata::cmip6_world(model = "CNRM-CM6-1", ssp = "585", 
                          res = 10, time = "2061-2080",
                          var = "bioc",
                          path = cmip6_path)

```

### Listing what you have

Use `list_geocache` to list the items you have stored.

```{r listing_simple}
# simple listing
list_geocache()
```

```{r listing_recurse}
# recursive listing
list_geocache(recursive = TRUE)
```

```{r list_sub_full}
# full names within a subdirectory
list_geocache(cmip6_path, full.names = TRUE, recursive = TRUE)
```


### About CMIP6

There is a lot of info about [CMPI6](https://esgf-node.llnl.gov/projects/cmip6/) with data distribution [here](https://www.worldclim.org/).  Some if it seems a bit opaque to the new user.  We have provided some functions to help navigate the data lingo.

```{r ssp}
knitr::kable(SSP_metadata())
```

```{r bioclim}
knitr::kable(CMPI6_bioclim_metadata())
```
